---
title: "Velo.com revisited assignment"
author: "Harichandana Gonuguntla"
date: "March-12,2023"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

<!-- Note:   -->

<!-- These instructions are commented out and will not display when you knit your RMarkdown document. -->

<!-- - Change the information in the yaml header above:  title, author, date. -->
<!-- - Make sure output is html_document. -->
<!-- - Once you are finished coding, **run each chunk individually to make sure there are no errors**.  (If necessary fix your code.) Once your code is error-free, click "knit" on the menu above. Your document should compile to HTML, provided that you have output set to "html_document." -->
<!-- - In the code chunk above ("setup") echo is set to TRUE.  This means that the code in your chunks will be displayed, along with the results, in your compiled document. -->

## Load packages and create empty data frame

```{r}
library(tidyverse)
v <- read_csv("velo.csv")

```

Note that `lm()` will automatically assign alphabetic factor levels to character predictors.  To make the regression coefficient for checkout_system easier to interpret in this tutorial we will manually change the factor levels.  This will ensure that the coefficient is, appropriately, positive.

```{r}

v <- v %>% 
  mutate(checkout_system = factor(checkout_system, 
                                  levels = c("old", "new"))) # or

# v$checkout_system <- factor(v$checkout_system, levels = c("old", "new"))


```

## Questions

Lightly comment your code and use pipes for readability.

Comment briefly on each of the questions, as directed.  Only the the final question requires a lengthier response.

The point of this case is to revisit velo.com  with a different methodology.  For reference, in the previous module you we obtained the following results from a t-test comparing the old and new systems:

```{r}
# T-test comparing old and new systems
t.test(filter(v, checkout_system == "new")$spent, 
       filter(v, checkout_system == "old")$spent)

# T-test comparing old and new systems for mobile users
t.test(filter(v, checkout_system == "new" & device == "mobile")$spent, 
       filter(v, checkout_system == "old" & device == "mobile")$spent)

```

### Q1

Fit a simple linear regression with `spent` as the target variable and `checkout_system` as the predictor. Based on the model summary, answer the following questions:

1. What is average spending for customers using 1) the new system and 2) the old system? This information can be extracted directly from the linear model output.

2. What is the difference in average customer spending between the new and old systems, and is the difference statistically significant at the p < .05 level?

3. Compare these results to those you obtained using a t-test in the last module (see the above output).

```{r}
lm(spent ~ checkout_system , data = v) 

# Regression equation:  spent = 2217.15 + 62.74*checkout_systemnew
# checkout_system is a binary 0/1 variable, so: spent = 2217.15 + 62.74*0 or spent = 2217.15 + 62.74*1.
# Simplifying: spent = 2217.15 (the intercept) or spent = 2217.15 + 62.74 (intercept + coefficient)

model <- lm(spent ~ checkout_system , data = v) # saving the model object as "model" for later reference
summary(model)

```

> Answer:  

>1. Average spending for the customers using old system is 2217.15$ and using new system it is equal to 2279.89$

>2. The difference between average customer spending between the new and old systems is 62.74$.( which is not very significant difference).If p<0.05,the result is statistically significant.But here p=0.1542(>0.05) which means the result is statistically not significant.

>3. The avearge customer spending in new and old checkout systems for mobile device using Linear regression model are 2279.89$ and 2217.15$ respectively. While the same values using t-test are 2279.89$ and 2217.148$ respectively.That means there is no difference in the mean earnings using the 2 methods.


### Q2

Fit a simple linear regression with `spent` as the target variable and `checkout_system` as the predictor, but include only customers who were using a mobile device. (Later you will learn how to fit this sort of regression using all the data, but for now subset the data to include only mobile users.) Answer these questions based on the model summary for this regression:

1. What is the difference in mobile customer spending between the new and old systems?

2. Is the difference statistically significant at the p < .05 level?

3. Compare these results to those you obtained using a t-test in the last module (see the above output).

Remember that the syntax for creating a simple linear regression model in R has the form:  `lm(y ~ x, data)`. This code will produce estimates of the model intercept and coefficients which you can use to assemble the regression equation:  y = intercept + coefficient * x.  To get more information about the model (including standard errors and p-values) use the generic `summary()` function. 


```{r}

lm(spent ~ checkout_system == "new", data = filter(v, device == "mobile")) %>% 
  summary

lm(spent ~ checkout_system == "old", data = filter(v, device == "mobile")) %>% 
  summary

t.test(filter(v, checkout_system == "old", device == "mobile")$spent,
       filter(v, checkout_system == "new", device == "mobile")$spent)

```

> Answer: 

>1. The difference in mobile customer spending between the new and old systems is 2174.92-2323 = 148.08$

>2. p< 0.05 result is statistically significant.But here p=0.016 (<0.05) which means the result is statistically significant, which means the difference in earnings is not by chance.

>3. The mobile customer spending in the new and old checkout systems using linear regression model and t-test are 2323$ and 2174.92$. That is the results are same in both the methods.


###  Q3

Using the summary of the regression model object from Q2, calculate a 95% confidence interval (CI) for the `checkout_system` coefficient using 1.96 as the critical value. Interpret the CI: What does it mean exactly?

```{r}
# upper CI:  
(148.08 + 1.96 * 61.98)

# lower CI
(148.08 - 1.96 * 61.98)

```

> Answer: A 95% CI that does not include 0 is equivalent to statistical significance: p < 0.05. From the above result, we can notice that the CI's do not include zero and hence the result is statistically significant.

### Q4

Based on the model output in Q2, along with the 95% CI you calculated in Q3, develop best and worst case revenue projections for the new checkout system.  What range of increased revenue might the company expect using the new checkout system (compared to the old system) and, based on this, does the coefficient estimate for `checkout_system`  have *practical* significance in your view? (Hint: multiply the lower and upper bounds of the CI you calculated in Q3 by an arbitrary number of customers.  That range, remember, is expressed in terms of average dollars per customer.)

A 95% CI for a coefficient represents the most likely range of values under repeated sampling.  Working with a range is advantageous.  Velo.com executives might well be interested in the low end of the range in order to assess the financial risks associated with implementing the new system.  Additionally, scaling up the average spending per customer gives you a way to convey the practical significance of using the new system. 

What do I mean by "scaling up"? For communication purposes it is often helpful to multiply averages, like the upper and lower bounds of the confidence interval, by some arbitrary scaling number (in this case, for example, 1000 or 5000 customers or whatever you think makes sense) so that a decision-maker can get a better sense of the overall impact to the business of a proposed change.  Think back to the Conley Fisheries case where you calculated that the maximum daily loss per boat was /$10,000.  By itself, that may seem like a sustainable loss. But remember there are 50 boats. So, really, to convey a complete sense of the risk the maximum per boat daily loss must be scaled up --- multiplied by --- the number of boats in the fleet. 

```{r}

#For New Checkout System
(148.08 + 2 * 61.98)*500  #Bestcase_projection_new
(148.08 - 2 * 61.98)*500  #Worstcase_projection_new

#For Old Checkout System
(-148.08 + 2 * 61.98)*500  #Bestcase_projection_old
(-148.08 - 2 * 61.98)*500  #Worstcase_projection_old

```

> Answer: Assuming 500 customers as the random scaling number,the best case revenue projection for the new checkout system is 136020$ and the worst case revenue projection is 12060$.From the above results it is clearly evident that the new checkout system gives profits(+ve output values) and the old checkout system gives losses(-ve output values) and the difference of earnings between these 2 systems has a huge practical significance. Hence the cofficient estimate for the checkout system has practical significance in my view.


### Q5

What course of action should Sarah recommend to the management at velo.com based on this additional analysis? Please incorporate your analytic results from above in fashioning an answer. In particular, cite the results from the new analysis, the simple linear regression, and discuss your interpretation of the range you developed in Question 4.

>Answer: Sarah has to recommend the management at velo.com to switch to the new checkout system from the old one. Her suggestion comes from the analysis she has performed and the outputs she has noticed using a simple linear regression model.The difference between mean earnings between old and new checkout systems using the simple linear regression model is 2279.89$ - 2217.15$ = 62.74$ .But when we include a parameter(type of device i.e mobile) the mean earnings between old and new checkout systems for mobile users using the simple linear regression model is 2323$ - 2174.92$ = 148.08$ .The p-value using above method without including mobile device as a filter is 0.1542 (>0.05) which says that the result is statistically insignificant. But when mobile devie type is considered the p-value is 0.0169 (<0.05 )and this means that the result is statistically significant and it is not by chance. Also the 95% CI values of mean earnings for mobile users are 269.5608$ and 26.5992$ .These CIs do not include 0 and this also states that the result is statistically significant. While statistical significance shows if an effect exists in a study, practical significance shows if the effect is large enough to be meaningful in the real world. In order to prove the practical significance in real world scenario, lets assume 500 customers as the random scaling number. With this, the best case revenue projection of mobile users for the new checkout system is 136020$ and the worst case revenue projection is 12060$ and for the old checkout system it is -136020$ and -12060$ respectively.So it is clearly evident that the new checkout system gives profits and the old one gives losses and the difference of earnings between these 2 systems has a huge practical significance. In consideration of all the above inferences from the simple linear regression model backed by strong data points to prove, Sarah can confidently suggest Velo management to switch to the new system in order to reap the benefits/profits it provides.


### Challenge


Read through the tutorial on the bootstrap (included in the lightbulb package).  Use the technique to estimate a confidence interval for `checkout_system` regression coefficient.  Does it agree with the confidence interval computed using the SEs for the coefficient from the regression model output?  Please show your code for your calculations.
